Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-22 cis-perfect-testing.tex

By definition (see \cpageref{perf-test:def:admissible}), we have, for any $i \in \set{D}$:
\begin{align}
p_{ia} =& \prob((B_j, E_j) \in \alpha_i \mid \vec{\theta}, i(j) = i) \\
\alpha_i =& \{ (b, e) : l_{j}^{(b)} \leq b \leq r_{j}^{(b)} \wedge l_{j}^{(e)} \leq e \leq r_{j}^{(e)}\}.
\intertext{Omitting the conditioning on $\vec{\theta}$ and $i(j) = i$, this gives:}
p_{ia}
=& \prob \left( l_{j}^{(b)} \leq B_{j} \leq r_{j}^{(b)}, l_{j}^{(e)} \leq E_{j} \leq r_{j}^{(e)} \right) \\
=& \prob \left( l_{j}^{(e)} \leq E_{j} \leq r_{j}^{(e)} \mid l_{j}^{(b)} \leq B_{j} \leq r_{j}^{(b)} \right) \\
   &\times\prob \left( l_{j}^{(b)} \leq B_{j} \leq r_{j}^{(b)} \right) \\
=& \sum_{b = l_{j}^{(l)}}^{r_{j}^{(b)}} \prob \left( l_{j}^{(e)} \leq E_{j} \leq r_{j}^{(e)} \mid B_{j} = b \right) \prob \left(B_{j} = b \right) \\
=& \sum_{b = l_{j}^{(b)}}^{r_{j}^{(b)}} \prob \left( l_{j}^{(e)} - b + 1 \leq D_{j} \leq r_{j}^{(e)} - b + 1 \right) \prob \left(B_{j} = b \right) &\text{by def of $D_{j}$} \\
=& \sum_{b = l_{j}^{(b)}}^{r_{j}^{(b)}} \left( S_{\vec{\theta}}(l_{j}^{(e)} - b + 1) - S_{\vec{\theta}}(r_{j}^{(e)} - b + 1) \right) \prob \left(B_{j} = b \right) &\text{by def of $S_{\vec{\theta}}$} \\
\propto& \sum_{b = l_{j}^{(b)}}^{r_{j}^{(b)}} \left( S_{\vec{\theta}}(l_{j}^{(e)} - b + 1) - S_{\vec{\theta}}(r_{j}^{(e)} - b + 1) \right)
\end{align}
under the assumption of uniform probability of infection time.
This is the standard form of the likelihood for doubly interval censored data without truncation~\autocite[e.g.][]{sunEmpirical}.

>>>>

By definition (see \cpageref{perf-test:def:admissible}), we have, for any $i \in \set{D}$:
\begin{align}
p_{ia} =& \prob((B_j, E_j) \in \alpha_i \mid \vec{\theta}, i(j) = i) \\
\alpha_i =& \{ (b, e) : l_{j}^{(b)} \leq b \leq r_{j}^{(b)} \wedge l_{j}^{(e)} \leq e \leq r_{j}^{(e)}\}.
\intertext{Omitting the conditioning on $\vec{\theta}$ and $i(j) = i$, this gives:}
p_{ia}
=& \prob \left( l_{j}^{(b)} \leq B_{j} \leq r_{j}^{(b)}, l_{j}^{(e)} \leq E_{j} \leq r_{j}^{(e)} \right) \\
=& \prob \left( l_{j}^{(e)} \leq E_{j} \leq r_{j}^{(e)} \mid l_{j}^{(b)} \leq B_{j} \leq r_{j}^{(b)} \right) \\
   &\times\prob \left( l_{j}^{(b)} \leq B_{j} \leq r_{j}^{(b)} \right) \\
=& \sum_{b = l_{j}^{(l)}}^{r_{j}^{(b)}} \prob \left( l_{j}^{(e)} \leq E_{j} \leq r_{j}^{(e)} \mid B_{j} = b \right) \prob \left(B_{j} = b \right) \\
=& \sum_{b = l_{j}^{(b)}}^{r_{j}^{(b)}} \prob \left( l_{j}^{(e)} - b + 1 \leq D_{j} \leq r_{j}^{(e)} - b + 1 \right) \prob \left(B_{j} = b \right) &\text{by def of $D_{j}$} \\
=& \sum_{b = l_{j}^{(b)}}^{r_{j}^{(b)}} \left( S_{\vec{\theta}}(l_{j}^{(e)} - b + 1) - S_{\vec{\theta}}(r_{j}^{(e)} - b + 2) \right) \prob \left(B_{j} = b \right) &\text{by def of $S_{\vec{\theta}}$} \\
\propto& \sum_{b = l_{j}^{(b)}}^{r_{j}^{(b)}} \left( S_{\vec{\theta}}(l_{j}^{(e)} - b + 1) - S_{\vec{\theta}}(r_{j}^{(e)} - b + 2) \right)
\end{align}
under the assumption of uniform probability of infection time.
This is the standard form of the likelihood for doubly interval censored data without truncation~\autocite[e.g.][]{sunEmpirical}.

### 2024-03-22 distributions.tex

\begin{landscape}
\section{Continuous distributions}
\begin{tabular}{llp{3.5cm}lll}
Name & Notation & Parameters & Mean & Variance & pdf\\
\hline \\
Normal & $\N(\mu, \sigma^2)$ & $\mu$ (mean)\newline $\sigma^2 > 0$ (variance) & $\mu$ & $\sigma^2$ & $\frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{(x-\mu)^2}{2\sigma^2}}$ \\
Lognormal & $\LN(\mu, \sigma^2)$ & $\mu$ (mean of the logarithm)\newline $\sigma^2 > 0$ (variance of the logarithm) & $e^{\mu + \frac{\sigma^2}{2}}$ & $(e^{\sigma^2} - 1)e^{2\mu + \sigma^2}$ & $\frac{1}{x\sigma\sqrt{2\pi}}e^{-\frac{(\ln(x)-\mu)^2}{2\sigma^2}}$ \\
Gamma & $\GamDist(\alpha, \beta)$ & $\alpha > 0$ (shape) \newline $\beta > 0$ (rate) & $\frac{\alpha}{\beta}$ & $\frac{\alpha}{\beta^2}$ & $\frac{\beta^\alpha}{\Gamma(\alpha)\theta}x^{\alpha-1}e^{-\beta x}$ \\
Beta & $\BetaDist(\alpha, \beta)$ & $\alpha > 0$ (shape)\newline $\beta > 0$ (shape) & $\frac{\alpha}{\alpha+\beta}$ & $\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}$ & $\frac{1}{B(\alpha,\beta)}x^{\alpha-1}(1-x)^{\beta-1}$ \\
Exponential & $\Exponential(\lambda)$ & $\lambda$ (rate) & $\frac{1}{\lambda}$ & $\frac{1}{\lambda^2}$ & $\lambda e^{-\lambda x}$ \\
LKJ* & $\LKJ(\eta)$ & $\eta > 0$ (shape) & $\matr{I}$ (identity matrix) & N/A & $C \det(\matr{x})^{\eta-1}$ \\
\end{tabular}

>>>>

\begin{landscape}
\section{Continuous distributions}
\begin{tabular}{llp{3.5cm}lll}
Name & Notation & Parameters & Mean & Variance & pdf\\
\hline \\
Normal & $\N(\mu, \sigma^2)$ & $\mu$ (mean)\newline $\sigma^2 > 0$ (variance) & $\mu$ & $\sigma^2$ & $\frac{1}{\sqrt{2\pi\sigma^2}}e^{-\frac{(x-\mu)^2}{2\sigma^2}}$ \\
\makecell{Multivariate \\ normal} & $\MNorm(\vec{\mu}, \matr{\Sigma})$ & $\mu \in \reals^k$ (mean vector)\newline $\Sigma \in \reals^{k \times k}$ (covariance matrix, positive semi-definite) & $\vec{\mu}$ & $\matr{\Sigma}$ & $(2 \pi)^{-k / 2} \operatorname{det}(\matr{\Sigma})^{-1 / 2} \exp \left(-\frac{1}{2}(\vec{x}-\vec{\mu})^T \matr{\Sigma}^{-1}(\vec{x}-\vec{\mu})\right)$ \\
Lognormal & $\LN(\mu, \sigma^2)$ & $\mu$ (mean of the logarithm)\newline $\sigma^2 > 0$ (variance of the logarithm) & $e^{\mu + \frac{\sigma^2}{2}}$ & $(e^{\sigma^2} - 1)e^{2\mu + \sigma^2}$ & $\frac{1}{x\sigma\sqrt{2\pi}}e^{-\frac{(\ln(x)-\mu)^2}{2\sigma^2}}$ \\
Gamma & $\GamDist(\alpha, \beta)$ & $\alpha > 0$ (shape) \newline $\beta > 0$ (rate) & $\frac{\alpha}{\beta}$ & $\frac{\alpha}{\beta^2}$ & $\frac{\beta^\alpha}{\Gamma(\alpha)\theta}x^{\alpha-1}e^{-\beta x}$ \\
Beta & $\BetaDist(\alpha, \beta)$ & $\alpha > 0$ (shape)\newline $\beta > 0$ (shape) & $\frac{\alpha}{\alpha+\beta}$ & $\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}$ & $\frac{1}{B(\alpha,\beta)}x^{\alpha-1}(1-x)^{\beta-1}$ \\
Exponential & $\Exponential(\lambda)$ & $\lambda$ (rate) & $\frac{1}{\lambda}$ & $\frac{1}{\lambda^2}$ & $\lambda e^{-\lambda x}$ \\
LKJ* & $\LKJ(\eta)$ & $\eta > 0$ (shape) & $\matr{I}$ (identity matrix) & N/A & $C \det(\matr{x})^{\eta-1}$ \\
\end{tabular}
